---
title: Flood Insurance Impact on Post-Flood Home Sales
author: Connor P. Jackson
date: "`r format(Sys.time(), '%B %d, %Y - %H:%M %Z')`"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
    number_sections: yes
    df_print: kable
  html_document:
    code_folding: show
    fig_caption: yes
    highlight: textmate
    number_sections: yes
    toc: yes
    toc_depth: 2
    toc_float: yes
    df_print: kable
  html_notebook:
    fig_caption: yes
    number_sections: yes
    df_print: kable
linestretch:
fontsize:
linkcolor: blue
urlcolor: blue
---

```{r setup, include=FALSE}
library(knitr)
library(rmarkdown)
options(knitr.kable.NA = '')
```

```{r initialize-data, include=FALSE}
source("merge_and_analysis.R")
```

Questions of interest:

1. Are homeowners more likely to sell following a flood event in their area (not looking at damage to specific homes)
2. Does holding a flood insurance policy mediate this effect (instrumenting for insurance takeup with exogenous policy shocks driving price changes)

Empirical method: event study, looking at flood events up to 3 years prior, and whether they held flood insurance in the year of the flood. Y variable is transaction probability in a given year. Currently running linear probability model. Might switch to probit. Instrumenting for flood insurance takeup. Data represents the set of homes in each census tract x flood zone x year of construction. Includes home set and year fixed effects. Will likely add (or swap in) census tract (or county) fixed effects for the spatial FEs. Clustering on the spacial fixed effects.

Coefficients of interest for each question:

1. `flood_eventTRUE`, `flood_L1TRUE`, `flood_L2TRUE`, `flood_L3TRUE`
2. `flood_eventTRUE:policy_prob`, `flood_L1TRUE:policy_prob_L1`, `flood_L2TRUE:policy_prob_L2`, `flood_L3TRUE:policy_prob_L3`


```{r event-study}
ols_res <- plm(transaction_prob ~ flood_event * policy_prob + flood_L1 * policy_prob_L1
                                  + flood_L2 * policy_prob_L2 + flood_L3 * policy_prob_L3
                                  | flood_event * reg_reform:adapted + flood_L1 * reg_reform_L1:adapted
                                    + flood_L2 * reg_reform_L2:adapted + flood_L3 * reg_reform_L3:adapted,
               data = tzy_panel[in_sample == TRUE], model = "within", index = c("panel_id", "year"), effect = "twoways")
summary(ols_res, vcov=vcovHC(ols_res, type="HC1", cluster="group"))
```


```{r first-stage}
# a rough test of the first stage (not the lagged versions)
testols <- lm(policy_prob ~ flood_event * reg_reform * adapted + year, data = tzy_panel[in_sample == TRUE])
summary(testols)
```


